rgbif: a package for working with species occurrence data in R

### 2015-10-05 manuscript.Rmd
deleted cache stuff, gitignoring now, updates to occurence sectin

The GBIF reqistry API services are spread across four major sets
of functions:

@@

GBIF provides two ways to get occurrence data: through the
/occurrence/search route (see `occ_search`), or via the
/occurrence/download route (many functions, see below).
`occ_search` is more appropriate for smaller data, while
occ_download*() functions are more appropriate for larger
data requests.

@@

The download API syntax is similar to the occurrence search API
in that the same parameters are used, but the syntas is quite different.
For example, in the download API you can do greater than searches,
whereas you can not do that in the occurrence search API.

```{r}
occ_download_get("0000066-140928181241064")
```

>>>>

The GBIF reqistry API services are spread across four sets
of functions:

@@

GBIF provides two ways to get occurrence data: through the
`/occurrence/search` route (see `occ_search`), or via the
`/occurrence/download` route (many functions, see below).
`occ_search()` is the main funtion for the search route,
and is more appropriate for smaller data, while
`occ_download*()` functions are more appropriate for larger
data requests.

Large is of course a subjective term. When you hit a "large dataset"
will depend primarily on the size of the your data request. GBIF imposes
for any given search a limit of 200,000 records in the search service,
after which point you can't download any more records for that search.
However, you can download more records for different searches.

We think the search service is still quite useful for many people
even given the 200,000 limit. For those that need more data, we have
created a similar interface in the `download_*()` functions, that should
be easy to use. Users should take note that using the download service
has a few extra steps to get data into R, but is straight-forward.

@@

The download API syntax is similar to the occurrence search API
in that the same parameters are used, but the way in which the query
is defined is different. For example, in the download API you can
do greater than searches, whereas you can not do that in the
occurrence search API. Thus, we can't make the query interace exactly
the same for both search and download functions.

Using the download service can be as few as three steps.

Request data download given a query. Here, we xxxx

```{r}
"xxx"
```

You can check on when the download is ready using the functions
`occ_download_list()` and `occ_download_meta()`. When it's ready
use `occ_download_get()` to download the dataset to your computer.

```{r}
(res <- occ_download_get("0000066-140928181241064", overwrite = TRUE))
```

What's printed out above is a very brief summary of what was downloaded,
the path to the file, and its size (in human readable form).

Next, read the data in to R using the function `occ_download_import()`.

```{r}
library("dplyr")
dat <- occ_download_import(res)
dat %>%
  select(gbifID, decimalLatitude, decimalLongitude)
```

### 2015-10-05 manuscript.tex
deleted cache stuff, gitignoring now, updates to occurence sectin

GBIF provides two ways to get occurrence data: through the
/occurrence/search route (see \texttt{occ\_search}), or via the
/occurrence/download route (many functions, see below).
\texttt{occ\_search} is more appropriate for smaller data, while
occ\_download*() functions are more appropriate for larger data
requests.

@@

The download API syntax is similar to the occurrence search API in that
the same parameters are used, but the syntas is quite different. For
example, in the download API you can do greater than searches, whereas
you can not do that in the occurrence search API.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{occ_download_get}\NormalTok{(}\StringTok{"0000066-140928181241064"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

>>>>

GBIF provides two ways to get occurrence data: through the
\texttt{/occurrence/search} route (see \texttt{occ\_search}), or via the
\texttt{/occurrence/download} route (many functions, see below).
\texttt{occ\_search()} is the main funtion for the search route, and is
more appropriate for smaller data, while \texttt{occ\_download*()}
functions are more appropriate for larger data requests.

Large is of course a subjective term. When you hit a ``large dataset''
will depend primarily on the size of the your data request. GBIF imposes
for any given search a limit of 200,000 records in the search service,
after which point you can't download any more records for that search.
However, you can download more records for different searches.

We think the search service is still quite useful for many people even
given the 200,000 limit. For those that need more data, we have created
a similar interface in the \texttt{download\_*()} functions, that should
be easy to use. Users should take note that using the download service
has a few extra steps to get data into R, but is straight-forward.

@@

The download API syntax is similar to the occurrence search API in that
the same parameters are used, but the way in which the query is defined
is different. For example, in the download API you can do greater than
searches, whereas you can not do that in the occurrence search API.
Thus, we can't make the query interace exactly the same for both search
and download functions.

Using the download service can be as few as three steps.

Request data download given a query. Here, we xxxx

\begin{Shaded}
\begin{Highlighting}[]
\StringTok{"xxx"}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] "xxx"
\end{verbatim}

You can check on when the download is ready using the functions
\texttt{occ\_download\_list()} and \texttt{occ\_download\_meta()}. When
it's ready use \texttt{occ\_download\_get()} to download the dataset to
your computer.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{(res <-}\StringTok{ }\KeywordTok{occ_download_get}\NormalTok{(}\StringTok{"0000066-140928181241064"}\NormalTok{, }\DataTypeTok{overwrite =} \OtherTok{TRUE}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

What's printed out above is a very brief summary of what was downloaded,
the path to the file, and its size (in human readable form).

Next, read the data in to R using the function
\texttt{occ\_download\_import()}.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"dplyr"}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{occ_download_import}\NormalTok{(res)}
\NormalTok{dat %>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(gbifID, decimalLatitude, decimalLongitude)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
      gbifID decimalLatitude decimalLongitude
1  657590544              NA               NA
2  657679551              NA               NA
3  657791316        37.70805        -118.4162
4  658180562              NA               NA
5  441881672              NA               NA
6  911596181              NA               NA
7   56454601              NA               NA
8  657848913              NA               NA
9  658187373              NA               NA
10 658279212        38.95917        -106.9892
..       ...             ...              ...
\end{verbatim}
